---
title: "Problem Set 6"
author: "Ethan Witkowski"
date: "Spring 2019"
output: 
  github_document:
    html_preview: true
    pandoc_args: --webtex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
<br>
<br>

#### 1)

```{r}
#read in data frame
data <- read.csv("nenana.csv")

# define variables
year <- data[,"year"]
date <- data[,"date"]

#Loess model
fit1 <- loess(date ~ year, span=.75)
plot(year, date)
lines(year, predict(fit1), col="red")

#Loess model 2 (smallest span)
fit2 <- loess(date ~ year, span=.25)
plot(year, date)
lines(year, predict(fit2), col="red")
```

There is evidence for a warming trend, indicated by earlier yearly ice break dates, especially from 1970 to 2019. When looking at the .75 span model, there appears to be an inflection point at 1970, where prior there is a relatively flat regression line, and after there is a more downward sloping regression line. When changing the span to .25, we see that there are changes in ice break date trends approximately every ten years, however, the general downward trend after 1970 persists.
<br>
<br>

#### 2a)

```{r}
#read in data frame
data2 <- read.csv("spam.csv")

#Define Variables
spam <- data2[,"spam"]
meeting <- data2[,"meeting"]
credit <- data2[,"credit"]

boxplot(meeting ~ spam)
```

There is ittle data available for spam messages, and all are near 0%. Skewed data for real messages.

```{r}
boxplot(credit ~ spam)
```

There is little data available for real messages, and all are near 0%. Skewed data for spam data.
<br>
<br>

#### 2b)

```{r}
fit3 <- glm(spam ~ meeting + credit + meeting*credit, family=binomial)
summary(fit3)
```

The interaction term is non-significant because the p-value is greater than .05. A possible explanation is that the sample size of messages with both the words "meeting" and "credit" is very small, or that the effect size is small.

```{r}
#Re-fit glm without interaction term
fit3 <- glm(spam ~ meeting + credit, family=binomial)
summary(fit3)
```
<br>
<br>

#### 2c)

The coefficient of credit is 13.65, and it is a statistically significant predictor.

Odds ratio = $\cfrac{e^{-.324+(13.65*2)}}{e^{-.324+(13.65*1)}} = e^{13.65} = 847460.92$
<br>
<br>

#### 2d)

The coefficient of meeting is -6.39. Meeting is a statistically significant predictor.

Odds ratio = $\cfrac{e^{-.324+(-6.39*2)}}{e^{-.324+(-6.39*1)}} = e^{-6.39} = .00168$

This is interpreted as a 1 percentage point increase in the prevalence of the word "meeting" predicts a multiplication of the odds of the message being spam by .00168.
<br>
<br>

#### 2e)

```{r}
#Create data
new <- as.data.frame(t(c(.2, 0)))
colnames(new) <- c("credit", "meeting")

predlogodds <- predict(fit3, newdata=new)
1 / (1 + exp(-predlogodds))
```
<br>
<br>

#### 2f)

```{r}
predlogodds <- predict(fit3)
predprobs <- 1 / (1 + exp(-predlogodds))
predspam <- predprobs >= .5

#Create table predicted spam, spam
table(predspam, spam)
```

Precentage of predicted spam that are actually spam:

$\cfrac{80}{87} = 92 \%$

Percentage of predicted non-spam that are actually non-spam:

$\cfrac{559}{913} = 61 \%$
<br>
<br>

#### 2g)

The spam filter is conservative because only 61% of the predicted non-spam were non-spam messages, while 92% of the predicted spam messages were actually spam messages.

We can change the aggressiveness of the spam filter by adjusting what the necessary predicted probability (predspam) is for the filter to consider it a spam message.