---
title: "Problem Set 5"
author: "Ethan Witkowski"
date: "Spring 2019"
output: 
  github_document:
    html_preview: true
    pandoc_args: --webtex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
<br>
<br>

#### 2a)

```{r}
homeprices <- read.csv("homeprices.csv")

# define variables
price <- homeprices[,"price"]
sqft <- homeprices[,"sqft"]
NE <- homeprices[,"NE"]

plot(price~sqft, col=NE+1)
#NE are light blue

# Linear model homes in NE and not in NE
fit1 <- abline(lm(price~sqft, subset=(NE==1)), col=1)
fit2 <- abline(lm(price~sqft, subset=(NE==0)), col=2)
```


The regression lines would have different slopes if NE and sqft had an interaction.

I do see evidence for an interaction, as the regression lines have different slopes.
<br>
<br>

#### 2b)

```{r}
fit3 <- lm(price ~ sqft + NE)
summary(fit3)

```

There is no evidence for a statistically significant interaction term, as the p-value for NE is .361.
<br>
<br>

#### 2c)

Yes, there is an outlier home with 3750 square feet.
<br>
<br>

#### 2d)

```{r}
homeprices <- homeprices[-79,]

# re-define variables
price <- homeprices[,"price"]
sqft <- homeprices[,"sqft"]
NE <- homeprices[,"NE"]

plot(price~sqft, col=NE+1)

# Linear model homes in NE and not in NE
fit1 <- abline(lm(price~sqft, subset=(NE==1)), col=1)
fit2 <- abline(lm(price~sqft, subset=(NE==0)), col=2)

fit3 <- lm(price ~ sqft + NE)
summary(fit3)
```

The slopes of the NE and non-NE regression lines are much closer.

The $\beta$ estimate of the NE interaction term is closer to 0 (now negative), and the p-value is much larger (.80).
<br>
<br>

#### 3a)

```{r}
congress <- read.csv("senate.csv")

# define variables
name <- congress[,"name"]
state <- congress[,"state"]
party <- congress[,"party"]
votes <- congress[,4:660]
```
<br>
<br>

#### 3b)

```{r}
pca <- prcomp(votes, center=T, scale=T)
PCscores <- pca$x
# set up plot with no points
plot(PCscores[,1], PCscores[,2], type="n")
# plot the names
text(PCscores[,1], PCscores[,2], label=name, cex=.4)
```
<br>
<br>

#### 3c)

```{r}
partycolor <- rep(NA, nrow(congress))
partycolor[party=="D"] <- "blue"
partycolor[party=="R"] <- "red"

pca <- prcomp(votes, center=T, scale=T)
PCscores <- pca$x
# set up plot with no points
plot(PCscores[,1], PCscores[,2], type="n")
# plot the names
text(PCscores[,1], PCscores[,2], label=name, cex=.4, col = partycolor)
```

The first principal compenent appears to represent the issue of immigration - as those on the left of the window oppose more immigration, while those on the right of the window support more immigration.
<br>
<br>

#### 3d)

These congresspeople are outliers because they only served partial terms.
<br>
<br>

#### 3e)

```{r}
congress <- congress[-c(42,43,44,53,54,63,64,65),]

name <- congress[,"name"]
state <- congress[,"state"]
party <- congress[,"party"]
votes <- congress[,4:660]

partycolor <- rep(NA, nrow(congress))
partycolor[party=="D"] <- "blue"
partycolor[party=="R"] <- "red"
```

The error occurs when attempting to perform the PCA because certain votes have 0 variance (all congresspeople voted the same way) -- this is likely a result from a vote like "March 6th is National Tree Day," where all the congresspeople would vote the same way.
<br>
<br>

#### 3f)

```{r}
#Find votes with 0 variance
#sort(apply(congress, 2, sd),
#      decreasing=T)

#Removing votes with 0 variance
congress <- congress[,-c(19,55,73,136,152,168,208,348,354,369,378,531,532,533,575)]

# define variables
name <- congress[,"name"]
state <- congress[,"state"]
party <- congress[,"party"]
votes <- congress[,4:645]

partycolor <- rep(NA, nrow(congress))
partycolor[party=="D"] <- "blue"
partycolor[party=="R"] <- "red"

#Run PCA without 0 variance columns/votes
pca <- prcomp(votes, center=T, scale=T)
PCscores <- pca$x
# set up plot with no points
plot(PCscores[,1], PCscores[,2], type="n")
# plot the names
text(PCscores[,1], PCscores[,2], label=name, cex=.4, col = partycolor)
```
<br>
<br>

#### 3g)

```{r}
summary(pca)
```

The first PCA accounts for 58.07% of the variance.

The second PCA accounts for 3.66% of variance by itself, 69.93% in cumulative with PCA1.

36 PC's are necessary to account for 90% of the variance in the dataset.
<br>
<br>

#### 3h)

The first PC appears to represent if the congress person supports pro-choice.

The second PC appears to represent voting on national debt, as there are many tea party members on the bottom left of the graph.

Murkowski and Collins are outliers, as they are moderate republicans as opposed to the more staunch republicans; for example, they are both pro-choice.
<br>
<br>

#### 3i)

Yes, King and Sanders, and they more closely resemble democrats.
<br>
<br>

#### 3j)

The democrats are more coherent in their voting.

It seems republicans are more spread on fiscal issues, as the tea party republicans are on the bottom left.

There also seems to be a trend in republicans where those who are more fiscally lenient are also more likely to vote pro-choice.


